### 2015-05-29 technical-background.tex
corrections

\begin{figure}
\includegraphics[width=\linewidth]{barco-ba411e-ipcore}
\caption{Barco Silex' AES IP Core}{There are two channels to comunicate with the IP core: the data and control paths~\cite{barco-ba411e}.}
\label{fig:barco-ba411e-ipcore}
\end{figure}

@@

A master interface has the advantage to offload data access operations to the device, but there is a ressource and area overhead to connect the DMA as well as an extra complexity to process its data.
A slave interface is usually used for sporadic data (\textit{e.g.} status, control flags), avoiding the overhead of a DMA.
As an example, the data path of Barco Silex' symmetric cryptography IP core is a master, as there is a lot of data to process, whilst the same data path of their asymmetric cryptography IP core is a slave, for the operands are larger and the computation takes more time.\newline{}

If a master interface with a DMA is used, the operating system buils scatterlists in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.

@@

Modern operating systems run on several levels of privilege.
Under Unix, the kernel runs under the highest level and it has no restriction, whereas applications run on the lowest level.
We call those two levels kernel space and user space, segragating the privilege and the memory between the two modes (\citet{tanenbaum2014} and \citet[chap. 2]{Corbet:2005:LDD:1209083}).
The figure~\ref{fig:user-kernel} shows this separation.
In order for the applications in the user space to reach the hardware, they need to go through the kernel and use the device drivers that are registered there.

\begin{figure}[ht]
\center
\input{user-kernel}
\caption{User and kernel space separation}{The applications and libraries from the the user space communicates with the kernel space via the system calls.}
\label{fig:user-kernel}
\end{figure}

@@

There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the status register fo the device, relentlessly asking if it finished its operations or tasks.
The first case is the cleanest and the most common: when the hardware device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will execute the interruption service routine (ISR) registered by the driver~\citep[chap. 10]{Corbet:2005:LDD:1209083}.
The second case is the easiest and is always guaranteed to work, but will monopolize the processir, loading it at 100\%, and avoiding any other task to be executed.
Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, \textit{i.e.} the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}

@@

By providing such an interface, the kernel applications\footnote{Used here as a misnomer to designate the code executed in kernel space.} can used cryptographic primitives without having to rely on their concrete implementation.
In order to add such an implementation, a kernel has to be developed and register itself with the Linux Crypto API upon its loading.
Those kernel modules can either be software implementation of some cryptogrpahic ciphers, or a device driver that will offload those operation to a dedicated hardware device.\newline{}

@@

Now rises the question of what and when authenticating.
\citet{Bellare2000} prooved that the most secure solution is to encrypt then compute the MAC from the ciphertext.
We will see in section~\ref{sec:theory-network} that if IPsec follows this recommandation, SSL/TLS does not and MAC first the plaintext then encrypt the message.

@@

Eventhough the signature algorithms are often more efficient than the usual MACs and the key pairs key stay in use for very long period of time, asymmetric cryptography can not supplant its symmetric counterpart.
The throughput rates of modern asymmetric schemes are still well under the symmetric schemes, plus the key sizes are much larger.
Whereas the most efficient search for the symmetric key (ignoring the weaknesses specific to certain modes) is an exhaustive search, all public-key systems are vulnerable to more efficient attacks (\textit{e.g.} factorization or discrete logarithm).

In practice, public-key cryptography is used to estalish a secure channel between two or more peers and exchange a shared secret key that will be used the encrypt the subsequent communications.

\subsubsection{RSA}
RSA is a public-key scheme proposed in 1978 by three MIT researchers (who gave it their name)~\cite{Rivest:1978:MOD:359340.359342}.
A few years later, they founded RSA Laboratories, which is now in charge of maintaining its standards, alongside many others, as the first Public-Key Cryptography Standards, \textit{aka} PKCS \#1.
The last version of the standard is the version 2.2~\cite{pkcs1} and is defined as a precise key generation protocol allowing encryption and decryption.
The keys can be generated by respecting a few steps:
\begin{enumerate}
	\item randomy choose two large primes $p$ and $q$;
	\item compute the modulus $n = p q$, and consequently we have \[\phi(n) = (p-1)(q-1)\]with $\phi(n)$ as the Euler function;
	\item randomly choose the public exponent $e \in ]1,\phi(n)[\ s.t.\ GCD(e,\phi(n)) = 1$;
	\item compute $d \in ]1,\phi(n)[\ s.t.\ e \cdot d \equiv 1 (mod\ \phi(n))$
\end{enumerate}

@@

The scheme is based on modular artihmetic, which is secure because of the factoring problem.
For an opponent to decrypt a ciphertext $c$ wihtout knowing the private key $(d, n)$, he would need to compute the $e^{th}$ roots of $c$, modulo $n$.
This operation is computationally hard, since an efficient solution is yet to be found for sufficiently large key size.

\subsubsection{Diffie-Hellman}
Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symmetric key during the following exchanges.
It uses the same kind of operation as RSA, that is modular exponentiation.
The protocol can be one of three types (\cite{rfc2631}, \cite{Frankel:2005:SGI:2206289}):
\begin{itemize}
	\item Static (DH): the actors use their authenticated certificate to compute the shared secret.
	\item Ephemeral (DHE): the actors create a new pair of public/private keys from which the secret key is derived.
	\item Anonymous: same as ephemeral, but without signing anything, hence not identifying neither of the actors, hence not providing any authentication. This mode is not advisable since it's vulnerable to main-in-the-middle attack.
\end{itemize}
A static scheme is easier to implement and requires much less operations, but using ephemeral keys is essential to ensure perfect forward secrecy.
Imagine that somehow, an opponent lays his hand on the shared secret.
If that secret has already been used, he can decipher all data transfered during past connections.
However, if the secret is new for every new connection, the compromission of the shared secret does not jeopardize past communications.
This is perfect forward secrecy: using a new key to protect the previous messages.

@@

If the server is Alice, it has to do at least one signature generation, one signature verification and two modular exponentiations.
Note that the client, B in our case, could have one signature less because RFC 5246~\cite{rfc5246} leave it as an optional feature, and the server would then have one verification less.
However, any sane configuration will have both actors signing their ephemeral public key.
If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of key.
We will see in chapter~\ref{chap:results} that while a 1024-bit key size is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit key sizes.
Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.

@@

% Present the TCP/IP layering
Computer networks are described by two models: the TCP/IP stack and the OSI model, each splitting the network workflow into more or less abstract layers.
The RFC 1122~\cite{rfc1122} defines the TCP/IP, and the international standard ISO/IEC 7498-1~\cite{ISOIEC7498} defines the OSI stack as in the table~\ref{tab:tcp-ip-stack}.
The main difference between the two is the application layer of the TCP/IP stack which corresponds to the three upper layers of the OSI model.
Some references, such as~\citet{tanenbaum2011}, conceptually split the TCP/IP link layer into an additional physical layer.

\begin{table}[ht]
\center
\begin{tabularx}{\textwidth}{|l|l|l|X|} \hline
\multicolumn{2}{|c|}{TCP/IP layering} & \multicolumn{2}{c|}{OSI model} \\ \hline
Layer & Protocols & Layer & Protocols \\ \hline
\multirow{3}{*}{Application} & \multirow{3}{*}{FTP, SSH} & Application & FTP \\ \cline{3-4}
 & & Presentation & ASCII, JPEG \\ \cline{3-4}
 & & Session & RPC, PAP \\ \hline
Transport & TCP, UDP & Transport & TCP, UDP \\ \hline
Internet & IP, ICMP, IPsec & Network & IP, ICMP, IPsec \\ \hline
\multirow{2}{*}{Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
 & & Physical & USB, DSL, IEEE 802.11 \\ \hline
\end{tabularx}
\caption{TCP/IP and OSI model comparison}{They are globaly the same, except for the application layer of the TCP/IP stack which merge togheter the three upper layers of the OSI model. Between parenthsis are examples of protocols resting on each layer.}
\label{tab:tcp-ip-stack}
\end{table}

@@

The present work will focus and SSL/TLS and IPsec, the figure~\ref{fig:ipsec-tls} showing their respective integration in the network stack.

@@

\subsection{SSL/TLS}
% Introduce SSL/TLS, talk about the protocol, the key exchange and stuff, but leaver OpenVPN for the 'implementation' chapter.
% Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.
Secure Socket Layer (SSL) and Transport Layer Security (TLS) are two protocols ensuring secure communications from the application layer, securing the transport layer.
The last version of SSL (version 3.0) dates from 1996~\cite{rfc6101} and has been found a critical vulnerability by Google in 2014~\cite{poodle}.
As a consequence, SSL 3.0 support is being dropped, and the IETF is working on a depracation of the standard~\cite{sslv3-diediedie-03}.

@@

\noindent The figure~\ref{fig:tls-handshake} shows the messages exchanged during the establishement of a connection between two peers.

@@

\begin{enumerate}
	\item Client: ClientHello. The client sends a series of specifications to the server, such as the protocol version and the supported cipher suites.
	\item Server: ServerHello and others. The server replies by setting the protocol and the cipher suite to use.
	It can send its own certificate and request the client's.
	In some cases, the server can already begin the key echange protocol (\textit{e.g.} Diffie-Hellman).
	\item Client: ClientKeyExchange and others. The client veryfies the validity of the server certificate, if any, and sends its own. If necessary, the client may have to change its cipher specification based on the constraints of the server, or request a new specification from the server.
	The client can also participate in the key exchange protocol.
	\item Server: Finished. If the client asked for a new cipher specification, the server can set a new one. Otherwise, the exchange is simply finished and the connection has been successfully established.
\end{enumerate}

Once the connection is established, the peers can exchange application data using the symmetric key negociated during the handshake.

\subsection{IPsec}
% IPsec as a protocol, strongswan come in the 'implementation' chapter.
IPsec is a protocol whose implementation is a modification of the IP stack in the kernel space.
It is a network/internet level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.

@@

\noindent The IKEv2 protocol uses four different types of exchanges to fulfill its role:
\begin{itemize}
	\item \texttt{IKE\_SA\_INIT}: The two peers agree on cryptographic parameters to use. Among these is a Diffie-Hellman shared secret from which are derived symmetric keys used for a special IKE SA.
	\item \texttt{IKE\_AUTH}: Once the peers are protected by the IKE SA, they can authenticate themselves to each other and add a first SA in the SA database (SADB).
	The protocol supports three types of authentication methods:
	\begin{itemize}
		\item Digital signature using a PKI.
		\item MAC using a pre-established secret key.
		\item Extensible Authentication Protocol (EAP) defined in~\citet{rfc3748}.
	\end{itemize}
	At this point, the connection is fully established and ready to use.
	\item \texttt{CREATE\_CHILD\_SA}: Used to create new SAs between the two peers. It may involve new generations of DH secrets to ensure perfect forward secrecy.
	\item \texttt{INFORMATIONAL}: Exchange of management information.
\end{itemize}
~\newline{}

@@

\noindent The ESP payload includes an optional IV and an optional Traffic Flow Confidentiality (TFC) padding.
The TFC padding allows to modify the length of the packet with dummy data, somewhat hiding the true nature of the payload.
It differs from the standard padding because its length is not limited to 255 bytes like the latter.

\noindent AH is the same as ESP, except is does not offer encryption features, only the authentication.
Since ESP can use a null cipher, it arguably offers the same features as AH.
The existance of AH is actually historical; when the protocol was developed in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
Having an authentication only counterparty allowed such product to be exported without encryption capabilities.\newline{}

IPsec can be deployed in two modes: 
\begin{itemize}
	\item Transport for end-to-end service and protecting the payload. IPsec examines the incomming payload, checks it has not been tampered with and pass it to the next layer.
	\item Tunnel protects the entire IP packet by encapsulating it into a new IP packet. As the destination contained in the new IP header can be different from the original, this mode can be used to manage gateways.
\end{itemize}
The figure~\ref{fig:ipsec-transport-tunnel} shows the space overhead imposed by AH and ESP in transport and tunnel modes.

>>>>

\begin{figure}
\includegraphics[width=\linewidth]{barco-ba411e-ipcore}
\caption{Barco Silex' AES IP Core}{There are two channels to communicate with the IP core: the data and control paths~\cite{barco-ba411e}.}
\label{fig:barco-ba411e-ipcore}
\end{figure}

@@

A master interface has the advantage to offload data access operations to the device, but there is a resource and area overhead to connect the DMA as well as an extra complexity to process its data.
A slave interface is usually used for sporadic data (\textit{e.g.} status, control flags), avoiding the overhead of a DMA.
As an example, the data path of Barco Silex' symmetric cryptography IP core is a master, as there is a lot of data to process, whilst the same data path of their asymmetric cryptography IP core is a slave, for the operands are larger and the computation takes more time.\newline{}

If a master interface with a DMA is used, the operating system builds scatterlists in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them any more, the DMA in now in charge to send them to the device where registers are ready to read the incoming data.
The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.

@@

Modern operating systems run on several levels of privilege.
Under Unix, the kernel runs under the highest level and it has no restriction, whereas applications run on the lowest level.
We call those two levels kernel space and user space, segregating the privilege and the memory between the two modes (\citet{tanenbaum2014} and \citet[chap. 2]{Corbet:2005:LDD:1209083}).
The figure~\ref{fig:user-kernel} shows this separation.
In order for the applications in the user space to reach the hardware, they need to go through the kernel and use the device drivers that are registered there.

\begin{figure}[ht]
\center
\input{user-kernel}
\caption{User and kernel space separation}{The applications and libraries from the user space communicates with the kernel space via the system calls.}
\label{fig:user-kernel}
\end{figure}

@@

There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the status register fo the device, relentlessly asking if it finished its operations or tasks.
The first case is the cleanest and the most common: when the hardware device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will execute the interruption service routine (ISR) registered by the driver~\citep[chap. 10]{Corbet:2005:LDD:1209083}.
The second case is the easiest and is always guaranteed to work, but will monopolize the processor, loading it at 100\%, and avoiding any other task to be executed.
Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, \textit{i.e.} the scheduler interrupts the running task and assigns the processor resources it used to another one.
Hence, systems with a lot of processes in need for CPU resources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}

@@

By providing such an interface, the kernel applications\footnote{Used here as a misnomer to designate the code executed in kernel space.} can used cryptographic primitives without having to rely on their concrete implementation.
In order to add such an implementation, a kernel has to be developed and register itself with the Linux Crypto API upon its loading.
Those kernel modules can either be software implementation of some cryptographic ciphers, or a device driver that will offload those operations to a dedicated hardware device.\newline{}

@@

Now rises the question of what and when authenticating.
\citet{Bellare2000} proved that the most secure solution is to encrypt then compute the MAC from the ciphertext.
We will see in section~\ref{sec:theory-network} that if IPsec follows this recommandation, SSL/TLS does not and MAC first the plaintext then encrypt the message.

@@

Even though the signature algorithms are often more efficient than the usual MACs and the key pairs key stay in use for very long period of time, asymmetric cryptography can not supplant its symmetric counterpart.
The throughput rates of modern asymmetric schemes are still well under the symmetric schemes, plus the key sizes are much larger.
Whereas the most efficient search for the symmetric key (ignoring the weaknesses specific to certain modes) is an exhaustive search, all public-key systems are vulnerable to more efficient attacks (\textit{e.g.} factorization or discrete logarithm).

In practice, public-key cryptography is used to establish a secure channel between two or more peers and exchange a shared secret key that will be used the encrypt the subsequent communications.

\subsubsection{RSA}
RSA is a public-key scheme proposed in 1978 by three MIT researchers (who gave it their name)~\cite{Rivest:1978:MOD:359340.359342}.
A few years later, they founded RSA Laboratories, which is now in charge of maintaining its standards, alongside many others, as the first Public-Key Cryptography Standards, \textit{aka} PKCS \#1.
The last version of the standard is the version 2.2~\cite{pkcs1} and is defined as a precise key generation protocol allowing encryption and decryption.
The keys can be generated by respecting a few steps:
\begin{enumerate}
	\item randomly choose two large primes $p$ and $q$;
	\item compute the modulus $n = p q$, and consequently we have \[\phi(n) = (p-1)(q-1)\]with $\phi(n)$ as the Euler function;
	\item randomly choose the public exponent $e \in ]1,\phi(n)[\ s.t.\ GCD(e,\phi(n)) = 1$;
	\item compute $d \in ]1,\phi(n)[\ s.t.\ e \cdot d \equiv 1 (mod\ \phi(n))$
\end{enumerate}

@@

The scheme is based on modular arithmetic, which is secure because of the factoring problem.
For an opponent to decrypt a ciphertext $c$ wihtout knowing the private key $(d, n)$, he would need to compute the $e^{th}$ roots of $c$, modulo $n$.
This operation is computationally hard, since an efficient solution is yet to be found for sufficiently large key size.

\subsubsection{Diffie-Hellman}
Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symmetric key during the following exchanges.
It uses the same kind of operation as RSA, that is modular exponentiation.
The protocol can be one of three types (\cite{rfc2631}, \cite{Frankel:2005:SGI:2206289}):
\begin{itemize}
	\item Static (DH): the actors use their authenticated certificate to compute the shared secret.
	\item Ephemeral (DHE): the actors create a new pair of public/private keys from which the secret key is derived.
	\item Anonymous: same as ephemeral, but without signing anything, hence not identifying neither of the actors, hence not providing any authentication. This mode is not advisable since it's vulnerable to main-in-the-middle attack.
\end{itemize}
A static scheme is easier to implement and requires much less operations, but using ephemeral keys is essential to ensure perfect forward secrecy.
Imagine that somehow, an opponent lays his hand on the shared secret.
If that secret has already been used, he can decipher all data transferred during past connections.
However, if the secret is new for every new connection, the compromise of the shared secret does not jeopardize past communications.
This is perfect forward secrecy: using a new key to protect the previous messages.

@@

If the server is Alice, it has to do at least one signature generation, one signature verification and two modular exponentiations.
Note that the client, B in our case, could have one signature less because RFC 5246~\cite{rfc5246} leave it as an optional feature, and the server would then have one verification less.
However, any sane configuration will have both actors signing their ephemeral public key.
If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of key.
We will see in chapter~\ref{chap:results} that while a 1024-bit key size is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit key sizes.
Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommendations since 2013~\cite{nist-sp800-131A}.

@@

% Present the TCP/IP layering
Computer networks are described by two models: the TCP/IP stack and the OSI model, each splitting the network work flow into more or less abstract layers.
The RFC 1122~\cite{rfc1122} defines the TCP/IP, and the international standard ISO/IEC 7498-1~\cite{ISOIEC7498} defines the OSI stack as in the table~\ref{tab:tcp-ip-stack}.
The main difference between the two is the application layer of the TCP/IP stack which corresponds to the three upper layers of the OSI model.
Some references, such as~\citet{tanenbaum2011}, conceptually split the TCP/IP link layer into an additional physical layer.

\begin{table}[ht]
\center
\begin{tabularx}{\textwidth}{|l|l|l|X|} \hline
\multicolumn{2}{|c|}{TCP/IP layering} & \multicolumn{2}{c|}{OSI model} \\ \hline
Layer & Protocols & Layer & Protocols \\ \hline
\multirow{3}{*}{Application} & \multirow{3}{*}{FTP, SSH} & Application & FTP \\ \cline{3-4}
 & & Presentation & ASCII, JPEG \\ \cline{3-4}
 & & Session & RPC, PAP \\ \hline
Transport & TCP, UDP & Transport & TCP, UDP \\ \hline
Internet & IP, ICMP, IPsec & Network & IP, ICMP, IPsec \\ \hline
\multirow{2}{*}{Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
 & & Physical & USB, DSL, IEEE 802.11 \\ \hline
\end{tabularx}
\caption{TCP/IP and OSI model comparison}{They are globally the same, except for the application layer of the TCP/IP stack which merge together the three upper layers of the OSI model. Between parenthesis are examples of protocols resting on each layer.}
\label{tab:tcp-ip-stack}
\end{table}

@@

The present work will focus and SSL/TLS and IPsec, the figure~\ref{fig:tls-ipsec} showing their respective integration in the network stack.

@@

\subsection{SSL/TLS}
% Introduce SSL/TLS, talk about the protocol, the key exchange and stuff, but leaver OpenVPN for the 'implementation' chapter.
% Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.
Secure Socket Layer (SSL) and Transport Layer Security (TLS) are two protocols ensuring secure communications from the application layer, securing the transport layer.
The last version of SSL (version 3.0) dates from 1996~\cite{rfc6101} and has been found a critical vulnerability by Google in 2014~\cite{poodle}.
As a consequence, SSL 3.0 support is being dropped, and the IETF is working on a deprecation of the standard~\cite{sslv3-diediedie-03}.

@@

\noindent The figure~\ref{fig:tls-handshake} shows the messages exchanged during the establishment of a connection between two peers.

@@

\begin{enumerate}
	\item Client: ClientHello. The client sends a series of specifications to the server, such as the protocol version and the supported cipher suites.
	\item Server: ServerHello and others. The server replies by setting the protocol and the cipher suite to use.
	It can send its own certificate and request the client's.
	In some cases, the server can already begin the key exchange protocol (\textit{e.g.} Diffie-Hellman).
	\item Client: ClientKeyExchange and others. The client verifies the validity of the server certificate, if any, and sends its own. If necessary, the client may have to change its cipher specification based on the constraints of the server, or request a new specification from the server.
	The client can also participate in the key exchange protocol.
	\item Server: Finished. If the client asked for a new cipher specification, the server can set a new one. Otherwise, the exchange is simply finished and the connection has been successfully established.
\end{enumerate}

Once the connection is established, the peers can exchange application data using the symmetric key negotiated during the handshake.

\subsection{IPsec}
% IPsec as a protocol, strongswan come in the 'implementation' chapter.
IPsec is a protocol whose implementation is a modification of the IP stack in the kernel space.
It is a network/internet level security; it examines incoming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.

@@

\noindent The IKEv2 protocol uses four different types of exchanges to fulfil its role:
\begin{itemize}
	\item \texttt{IKE\_SA\_INIT}: The two peers agree on cryptographic parameters to use. Among these is a Diffie-Hellman shared secret from which are derived symmetric keys used for a special IKE SA.
	\item \texttt{IKE\_AUTH}: Once the peers are protected by the IKE SA, they can authenticate themselves to each other and add a first SA in the SA database (SADB).
	The protocol supports three types of authentication methods:
	\begin{itemize}
		\item Digital signature using a PKI.
		\item MAC using a pre-established secret key.
		\item Extensible Authentication Protocol (EAP) defined in~\citet{rfc3748}.
	\end{itemize}
	At this point, the connection is fully established and ready to use.
	\item \texttt{CREATE\_CHILD\_SA}: Used to create new SAs between the two peers. It may involve new generations of DH secrets to ensure perfect forward secrecy.
	\item \texttt{INFORMATIONAL}: Exchange of management information.
\end{itemize}
~\newline{}

@@

\noindent The ESP payload includes an optional IV and an optional Traffic Flow Confidentiality (TFC) padding.
The TFC padding allows modifying the length of the packet with dummy data, somewhat hiding the true nature of the payload.
It differs from the standard padding because its length is not limited to 255 bytes like the latter.

\noindent AH is the same as ESP, except is does not offer encryption features, only the authentication.
Since ESP can use a null cipher, it arguably offers the same features as AH.
The existence of AH is actually historical; when the protocol was developed in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
Having an authentication only counterparty allowed such product to be exported without encryption capabilities.\newline{}

IPsec can be deployed in two modes: 
\begin{itemize}
	\item Transport for end-to-end service and protecting the payload. IPsec examines the incoming payload, checks it has not been tampered with and pass it to the next layer.
	\item Tunnel protects the entire IP packet by encapsulating it into a new IP packet. As the destination contained in the new IP header can be different from the original, this mode can be used to manage gateways.
\end{itemize}
The figure~\ref{fig:ipsec-transport-tunnel} shows the space overhead imposed by AH and ESP in transport and tunnel modes.

### 2015-05-29 test-proto.tex
corrections

\section{Experimental setup}
The experimental environment is built around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.
Both are linked toghether through a network capped by 100Mbps switches.
Both stations have gigabits ethernet interface and could hence be directly connected to each other, but in that case the communication would be limited by the I/O transfers of the storage units -- a hard drive disk in one case, an micro-SD card in the second -- on which we can not depend to set a constant throughput limitation, as it is highly influenced by the data block size and general health of the support.

\subsection{x86 host}
The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the developement and testings.

@@

\noindent The user should also be aware that if Streamline if a very powerfull and versatile tool, it is not reliable at very high resolution.
The data should not be analyzed in windows smaller than a millisecond.

@@

The table~\ref{tab:software-version} summarizes the softwares used and tjeir respective version.

@@

\noindent As AES mode CBC and SHA-2 are widely used, they will be used with a serious security paramter using keys of 256-bit.
Some test case will also experiment some implementation of AES-GCM to compare its performance and justify the need for its offloading in hardware.

@@

All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the perfomance could only be better.

\subsection{TLS Connections}
This benchmark is done only with OpenVPN 2.3.6.
Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPsec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
As soon as the public key operations can be plugged into the Linux Crypto API, this use case should however be tested.

For this use case, the board is configured in server mode, so that it can accept connections from any client.
The server does not connect with a specific client and waits for incomming connection requests (see listing~\ref{list:openvpn-server}).
The virtual machine will then execute then clients in parallel using the script~\ref{list:openvpn-client-script}.
The only option differentiating the clients is their IP address and port number.
Otherwise, all the clients share the same basic configuration file (see listing~\ref{list:openvpn-client}), which tell them to renogociate a new connection every second.
Hence, if a connection could be made with no delay and if the processes scheduling were ideal, the server would have to address 600 connections per minute.

@@

The security parameter tested is a standard TLS-DHE-RSA, hence forcing the peers the renogociate a new shared secret and ephemeral Diffie-Hellman parameters at each connection attempt.

@@

\subsection{File transfer}
The file transfered is an uncompressed block of 128MB of random data generated using the following command:
\begin{lstlisting}[language=bash]
  $ head -c $((1024*1024*128)) /dev/urandom > heavy.file
\end{lstlisting}~\newline{}

For OpenSSH, the commmand \texttt{scp} will be used to transfer the data securely over an SSH tunnel.
As for OpenVPN and IPsec, a tunnel will be established beforehand, and then the simple \texttt{ftp} command will allow the client to fetch the file on the server.
In order to use custom temporary security configuration with \texttt{scp}, the following command can be used:
\begin{lstlisting}[language=bash]
  $ ./scp -S /path/to/patched_openssh/bin/ssh -o Ciphers=aes256-cbc -o MACs=none@barco.com root@150.158.232.241:heavy.file .
\end{lstlisting}~\newline{}

All the results on the throughput and CPU usage take into account only the file transfer.
The server authentication, the connection establishement and any other security negociation is ignored in this particular test case.

>>>>

\section{Experimental setup}
The experimental environment is built around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.
Both are linked together through a network capped by 100Mbps switches.
Both stations have gigabits ethernet interface and could hence be directly connected to each other, but in that case the communication would be limited by the I/O transfers of the storage units -- a hard drive disk in one case, an micro-SD card in the second -- on which we can not depend to set a constant throughput limitation, as it is highly influenced by the data block size and general health of the support.

\subsection{x86 host}
The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the development and testings.

@@

\noindent The user should also be aware that if Streamline if a very powerful and versatile tool, it is not reliable at very high resolution.
The data should not be analysed in windows smaller than a millisecond.

@@

The table~\ref{tab:software-version} summarizes the software used and their respective version.

@@

\noindent As AES mode CBC and SHA-2 widely are used, they will be used with a serious security parameter using keys of 256-bit.
Some test case will also experiment some implementation of AES-GCM to compare its performance and justify the need for its offloading in hardware.

@@

All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the performance could only be better.

\subsection{TLS Connections}
This benchmark is done only with OpenVPN 2.3.6.
Since there is no standard support for those operations in the kernel yet, it would not have made sense to use IPsec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
As soon as the public key operations can be plugged into the Linux Crypto API, this use case should however be tested.

For this use case, the board is configured in server mode, so that it can accept connections from any client.
The server does not connect with a specific client and waits for incoming connection requests (see listing~\ref{list:openvpn-server}).
The virtual machine will then execute then clients in parallel using the script~\ref{list:openvpn-client-script}.
The only option differentiating the clients is their IP address and port number.
Otherwise, all the clients share the same basic configuration file (see listing~\ref{list:openvpn-client}), which tell them to renogotiate a new connection every second.
Hence, if a connection could be made with no delay and if the processes scheduling were ideal, the server would have to address 600 connections per minute.

@@

The security parameter tested is a standard TLS-DHE-RSA, hence forcing the peers the renogotiate a new shared secret and ephemeral Diffie-Hellman parameters at each connection attempt.

@@

\subsection{File transfer}
The file transferred is an uncompressed block of 128MB of random data generated using the following command:
\begin{lstlisting}[language=bash]
  $ head -c $((1024*1024*128)) /dev/urandom > heavy.file
\end{lstlisting}~\newline{}

For OpenSSH, the command \texttt{scp} will be used to transfer the data securely over an SSH tunnel.
As for OpenVPN and IPsec, a tunnel will be established beforehand, and then the simple \texttt{ftp} command will allow the client to fetch the file on the server.
In order to use custom temporary security configuration with \texttt{scp}, the following command can be used:
\begin{lstlisting}[language=bash]
  $ ./scp -S /path/to/patched_openssh/bin/ssh -o Ciphers=aes256-cbc -o MACs=none@barco.com root@150.158.232.241:heavy.file .
\end{lstlisting}~\newline{}

All the results on the throughput and CPU usage take into account only the file transfer.
The server authentication, the connection establishment and any other security negotiation is ignored in this particular test case.
